---
title: "Check-in 2"
author: "Shubhu Shahade"
date: "2023-08-28"
output: html_document
---

You should edit this .Rmd using RStudio, then click *Render* in the menu bar of the Source window (above the text of this .Rmd). Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit!

### Collaborators

INSERT NAMES OF ANY COLLABORATORS

## A. Functions

Work with one or two other students,

**1.** In words, what does the following line of code do?

```{r, eval = T}
sample(1:6, 1)
```

It prints a random number between 1 and 6.

**2.** Run the line of code in RStudio. What do you get? What did your groupmates get? What happens to the result each time you knit the document?

6 was printed as result. A different outcome between 1-6 is printed every time I knit the document.  

**3.** Below, we introduce the `set.seed` function. What do the two lines of code below do? What does the `456` mean? What happens to the result each time you knit the document?

```{r, eval = T}
set.seed(456)
sample(1:6, 1)
```

<!-- REPLACE THIS LINE WITH YOUR ANSWER --> The number 5 is printed each time I knit the document. It is because of the set.seed function. When we use the function, we always get a fixed value as output no matter what the input is. 456 doesn't really mean anything here. It is just a constant value that is being passed as a parameter over here.  

## B. Vectors and Matrices

**4.** Write a line of code to generate a vector of length 100 whose entries are 1, 2, ..., 100.

```{r, eval = T}
c(seq(1, 100))
```

**5.** Write a line of code to generate the following matrix:

`1 2 3`

`4 5 6`

`7 8 9`

```{r, eval = T}
 my_matrix <- matrix(1:9, nrow = 3, byrow = TRUE)
my_matrix
```
