---
title: "Check-in 3"
author: "Shubhu Shahade"
date: "2023-09-11"
format: 
  html:
    code-tools: true
    code-summary: "Code"
---

Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit! In addition, to receive full credit, your code output and plots must be correctly formatted.

### Collaborators

INSERT NAMES OF ANY COLLABORATORS

## A. Flights Data

**1.**  For this lab, we'll be using the `flights` data. You can load this data using the following code:

```{r}
#install.packages("dplyr")
library(dplyr)
library(nycflights13)
data(flights)
flights
```

**2.** Use the `mutate()` function to turn `origin` into a factor.

```{r}
# YOUR CODE HERE

# Using mutate to convert a column to a factor
flights <- flights %>%
  mutate(flights$origin == as.factor(flights$origin))

flights
```

**3.** Compute the mean `arr_delay` for each origin airport. Which airport has the longest delays on average? You can drop missing rows.

```{r}
# YOUR CODE HERE
origin_mean_arr_delay <- aggregate(flights$arr_delay ~ flights$origin, data = flights, FUN = mean, na.rm = TRUE)
origin_mean_arr_delay
```

<!--# YOUR ANSWER/EXPLANATION HERE --> EWR airport has the longest delays on average 


**4.** Use the `mutate()` function to turn `month` into a factor.

```{r}
# YOUR CODE HERE
flights <- flights %>%
  mutate(flights$month == as.factor(flights$month))

flights
```

<!--# YOUR ANSWER/EXPLANATION HERE -->


**5.** Compute the mean `arr_delay` by month. What do you notice?

```{r}
# YOUR CODE HERE
month_mean_arr_delay <- aggregate(flights$arr_delay ~ flights$month, data = flights, FUN = mean, na.rm = TRUE)
month_mean_arr_delay
```

<!--# YOUR ANSWER/EXPLANATION HERE --> Maximum delays occur during the months of June, July, December and April.There's negative delay during September and October. 


**6.** Compute the mean `arr_delay` by month AND airport. What do you notice?

```{r}
# YOUR CODE HERE
mean_arr_delay <- aggregate(flights$arr_delay ~ flights$month + flights$origin, data = flights, FUN = mean, na.rm = TRUE)
mean_arr_delay
```

<!--# YOUR ANSWER/EXPLANATION HERE --> The pattern is a bit consistent for all 3 airports. The all have a lot of delays in the last month of the year and during the months of June and July. EWR has significant delays running from December- April and then again in June and July. 

