---
title: "Check-in 4"
author: "Shubhu Shahade"
date: "2023-09-18"
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit! In addition, to receive full credit, your code output and plots must be correctly formatted.

### Collaborators

INSERT NAMES OF ANY COLLABORATORS

## A. Housing units data

```{r, warning = F, message = F}
library(dplyr)
library(tidyverse)
library(tidyr)
library(knitr)
housing_units <- readr::read_csv("https://www2.census.gov/programs-surveys/popest/datasets/2010-2020/housing/HU-EST2020_ALL.csv")
US_housing_units <- housing_units %>%
  filter(SUMLEV == "010")

US_housing_units
```

**1**. Generate a line plot of the number of housing units in the entire United States from 2010 to 2020.
```{r, warning = F, message = F}

US_housing_units<- US_housing_units |> 
  select(-HUESTIMATE042020) |>
  pivot_longer(contains("HUESTIMATE20"), 
               names_to = "Year",
               values_to = "HousingUnits") |>
  mutate(Year = as.numeric(stringr::str_sub(Year, start = 11)))
ggplot(data = US_housing_units, aes(x =Year, y =HousingUnits)) + 
  geom_line()
```


**2.** Generate a line plot of the number of housing units in each of the four regions (Midwest, Northeast, South, and West) from 2010 to 2020, using a different color for each region.
```{r, warning = F, message = F}
region_housing_units <- housing_units %>%
  filter(SUMLEV == "020")
region_housing_units

region_housing_units<- region_housing_units |> 
  select(-HUESTIMATE042020) |>
  pivot_longer(contains("HUESTIMATE20"), 
               names_to = "Year",
               values_to = "HousingUnits") |>
  mutate(Year = as.numeric(stringr::str_sub(Year, start = 11)))
ggplot(data = region_housing_units, aes(x =Year, y =HousingUnits, color = REGNAME)) + 
  geom_line()
```