---
title: "Check-in 8"
author: "Shubhu Shahade"
date: "2023-10-30"
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit! In addition, to receive full credit, your code output and plots must be correctly formatted.

### Collaborators
Quoc Luong Huynh

INSERT NAMES OF ANY COLLABORATORS

1. Write a function `sum_for_loop()`, that takes as input two vectors of length $n$, `x` and `y`, and returns the vector sum of `x` and `y` using a for loop of $n$ steps where the $i$th step involves adding `x[i]` and `y[i]`, Write a function `sum_vectorized()` that takes as input two vectors of length $n$, `x` and `y`, and computes the vector sum using the vectorized operator `x+y`.
```{r}
sum_for_loop <- function(x, y) {
  n <- length(x)
  if (length(y) != n) {
    stop("Input vectors must be of the same length")
  }
  result <- numeric(n)  # Initialize the result vector

  for (i in 1:n) {
    result[i] <- x[i] + y[i]
  }
  return(result)
}

sum_vectorized <- function(x, y) {
  if (length(x) != length(y)) {
    stop("Input vectors must be of the same length")
  }
  result <- x + y  
  return(result)
}

# Example data
x <- c(1, 2, 3, 4, 5)
y <- c(5, 4, 3, 2, 1)

# Using sum_for_loop
result_for_loop <- sum_for_loop(x, y)
print(result_for_loop)

# Using sum_vectorized
result_vectorized <- sum_vectorized(x, y)
print(result_vectorized)
```


2. Compare the speed of your two functions using one of the approaches discussed in class today.
```{r}
start_time_sum_vectorized <- Sys.time()
sum_vectorized(x,y)
print(Sys.time() - start_time_sum_vectorized)
```

```{r}
start_time_for_loop <- Sys.time()
sum_vectorized(x,y)
print(Sys.time() - start_time_for_loop)
```

From the time difference, we can see that sum_vectorized function is faster approach. However, the speed of the functions keeps varying upon every run. However, the vectorized sum function was generally faster. 