---
title: "Check-in 9"
author: "Shubhu Shahade"
date: "2023-11-08"
date-format: "[Due] MMMM DD, YYYY"
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit! In addition, to receive full credit, your code output and plots must be correctly formatted.

### Collaborators

INSERT NAMES OF ANY COLLABORATORS

1. (Devore Section 7.2 #5) Assume that the helium porosity (in percentage) of coal samples taken from any particular seam is normally distributed with true standard deviation .75. Compute a 90% CI for the true average porosity of a certain seam if the average porosity for 20 specimens from the seam was 4.85.
```{r}
# Given values
sample_mean <- 4.85
sample_size <- 20
true_standard_deviation <- 0.75
confidence_level <- 0.90


df <- sample_size - 1

margin_of_error <- qt((1 + confidence_level) / 2, df) * (true_standard_deviation / sqrt(sample_size))
confidence_interval <- c(sample_mean - margin_of_error, sample_mean + margin_of_error)
confidence_interval
```

2. The `trees` data set provides measurements for a sample of 31 felled black cherry trees. Based on this sample, create a 95% confidence interval for the population mean girth. You may use the command `data(trees)` to download this data.
```{r}
data(trees)
#head(trees)

# Given values
sample_mean <- mean(trees$Girth)
sample_size <- length(trees$Girth)
confidence_level <- 0.95 

# Degrees of freedom for a t-distribution
df <- sample_size - 1

sample_sd <- sd(trees$Girth)
margin_of_error <- qt((1 + confidence_level) / 2, df) * (sample_sd / sqrt(sample_size))
confidence_interval <- c(sample_mean - margin_of_error, sample_mean + margin_of_error)
confidence_interval
```