---
title: "Check-in 7"
author: "Shubhu Shahade"
date: "2023-10-23"
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit! In addition, to receive full credit, your code output and plots must be correctly formatted.

### Collaborators

INSERT NAMES OF ANY COLLABORATORS

Use the following code to create a list of four matrices:

```{r, echo = T, eval = T}
set.seed(100)
matrix_list <- list(
  A = diag(5),
  B = matrix(rnorm(9), nrow = 3, ncol = 3),
  C = matrix(c(1, 2, 3, 4), nrow = 2, ncol = 2),
  D = diag(c(1:5))
)
```

1. Use the `lapply` function to create a list of length four containing the inverse of these four matrices.
```{r, echo = T, eval = T}
inverse_matrix_list <- lapply(matrix_list, function(mat) solve(mat))
inverse_matrix_list

```

2. Use the `sapply` function to create a vector of length four containing the determinants of these four matrices.
```{r, echo = T, eval = T}
determinants <- sapply(matrix_list, function(mat) det(mat))
determinants

```