---
title: "Lab 3"
author: "SHUBHU SHAHADE"
date: "2023-09-06"
output: html_document
---

You should edit this .Rmd using RStudio, then click *Render* in the menu bar of the Source window (above the text of this .Rmd). Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit!

### Collaborators

INSERT NAMES OF ANY COLLABORATORS

```{r}
# LOAD ANY RELEVANT PACKAGES HERE
#install.packages("readr")
#install.packages("curl")
```

## A. Weather Forecast Data

1.  For this lab, we'll be using data on weather forecasts gathered by student at Saint Louis University. You can read about the dataset [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-12-20). Download the weather forecasts data using the following code:

```{r}
weather_forecasts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-20/weather_forecasts.csv')
weather_forecasts
```

2.  How many rows are in this dataset? How many columns?

```{r}
# YOUR CODE HERE
dim(weather_forecasts)
```

<!--# YOUR ANSWER/EXPLANATION HERE --> 651968 rows and 10 columns in the dataset 

3.  How many cities are represented in this dataset?

```{r}
# YOUR CODE HERE
city_count <- length(unique(weather_forecasts$city))
city_count
```

<!--# YOUR ANSWER/EXPLANATION HERE --> 160 cities in the dataset

4.  Create a new data frame containing only the forecasts for San Jose. You may have to explore the values for the `city` variable.

```{r}
# YOUR CODE HERE
sj_weather_forecasts <- weather_forecasts[weather_forecasts$city == "SAN_JOSE", ]
sj_weather_forecasts
```

<!--# YOUR ANSWER/EXPLANATION HERE --> 

5. Compute the mean absolute error between `observed_temp` and `forecast_temp` for San Jose.

```{r}
# YOUR CODE HERE
sj_weather_forecasts <- sj_weather_forecasts[complete.cases(sj_weather_forecasts), ]
absolute_differences <- abs(sj_weather_forecasts$observed_temp - sj_weather_forecasts$forecast_temp)
mean_abs_diff <- mean(absolute_differences)
mean_abs_diff
```

<!--# YOUR ANSWER/EXPLANATION HERE --> The difference in mean absolute error between observed and forecast temperature is 2.171611

6. Compute the mean absolute error between `observed_temp` and `forecast_temp` for San Jose using only forecasts made 48 hours in advance.

```{r}
# YOUR CODE HERE
forty_eight_hr_forecast <- sj_weather_forecasts[sj_weather_forecasts$forecast_hours_before == 48, ]
forty_eight_hr_forecast
absolute_differences <- abs(forty_eight_hr_forecast$observed_temp - forty_eight_hr_forecast$forecast_temp)
mean_abs_diff <- mean(absolute_differences)
mean_abs_diff
```

<!--# YOUR ANSWER/EXPLANATION HERE -->The difference in mean absolute error between observed and forecast temperature for 48 hrs in advance is 2.264327

7. Compute the mean absolute error between `observed_temp` and `forecast_temp` for San Jose using only forecasts made 12 hours in advance.

```{r}
# YOUR CODE HERE
twelve_hr_forecast <- sj_weather_forecasts[sj_weather_forecasts$forecast_hours_before == 12, ]
absolute_differences <- abs(twelve_hr_forecast$observed_temp - twelve_hr_forecast$forecast_temp)
mean_abs_diff <- mean(absolute_differences)
mean_abs_diff
```

<!--# YOUR ANSWER/EXPLANATION HERE --> The difference in mean absolute error between observed and forecast temperature for 12 hrs in advance is 2.057737

8. Compare your answers to 6 and 7. What do you notice? 

<!--# YOUR ANSWER/EXPLANATION HERE --> I can notice that the difference in mean absolute errors is less for the forecast made 12 hrs in advance than when compared to the forecast done 48 hrs in advance. 