---
title: "Lab 7"
author: "Shubhu Shahade"
date: "2023-10-18"
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit! In addition, to receive full credit, your code output and plots must be correctly formatted.

### Collaborators

```{r}
# LOAD ANY RELEVANT PACKAGES HERE
library(dplyr)
library(nycflights13)
```

## A. Skewness and Kurtosis

Skewness describes how asymmetrical the distribution of a numerical variable is around its mean. Given observations $x_1,\ldots, x_n$, we can calculate the sample skewness $s$ of a variable using the following formula:

$$s = \frac{\frac{1}{n}\sum\limits_{i=1}^n(x_i-\overline{x})^3}{\left[\frac{1}{n}\sum\limits_{i=1}^n(x_i-\overline{x})^2\right]^{3/2}}$$
Kurtosis is a measure of the "tailedness" of the distribution of a numerical variable is around its mean. Higher values of kurtosis indicate more extreme outliers. Given observations $x_1,\ldots, x_n$, we can calculate the sample kurtosis $k$ of a variable using the following formula:

$$k = \frac{\frac{1}{n}\sum\limits_{i=1}^n(x_i-\overline{x})^4}{\left[\frac{1}{n}\sum\limits_{i=1}^n(x_i-\overline{x})^2\right]^{2}}-3$$

1. Write a function `skewness()` that takes as input a numeric vector `x` and returns the sample skewness. There are functions in R that compute skewness, but you cannot use any of them--write your own implementation.
```{r}
skewness <- function(x) {
  n <- length(x)
  if (n < 3) {
    stop("Sample size should be at least 3 for skewness calculation.")
  }

  x_bar <- mean(x)

  numerator <- sum((x - x_bar)^3) / n
  denominator <- (sum((x - x_bar)^2) / n)^(3/2)

  s <- numerator / denominator
  
  return(s)
}

x <- c(1, 2, 3, 4, 5)
skew <- skewness(x)
print(skew)

```

2. Write a function `kurtosis()` that takes as input a numeric vector `x` and returns the sample skewness. There are functions in R that compute kurtosis, but you cannot use any of them--write your own implementation.
```{r}
kurtosis <- function(x) {
  n <- length(x)
  if (n < 4) {
    stop("Sample size must be at least 4 to calculate kurtosis")
  }
  
  x_bar <- mean(x)
  numerator <- sum((x - x_bar)^4) / n
  denominator <- (sum((x - x_bar)^2) / n)^2
    
                
  
  k <- (numerator / denominator) - 3
  return(k)
}

kurt <- kurtosis(x)
print(kurt)

```

3. Use your function to compute the sample skewness of the `arr_delay` variable in the `flights` dataset contained in the `nycflights13` package.
```{r}

data(flights)
flights
```

```{r}
flights <- subset(flights, !is.na(arr_delay))
skewness_result <- skewness(flights$arr_delay)
skewness_result

```


4. Use your function to compute the sample kurtosis of the `arr_delay` variable in the `flights` dataset contained in the `nycflights13` package.
```{r}
kurtosis_result <- kurtosis(flights$arr_delay)

print(kurtosis_result)

```
## B. Finding an error

Suppose you have two teams of runners participating in a 5k. We wish to write a function that takes as input two vectors representing the times of the runners in each team and returns a list of two vectors representing the ranks of each team's runners.

For example, if the first team's times are `c(16.8, 21.2, 19.1)` and the second team's times are `c(17.2, 18.1, 20.0)`, the function should return `c(1, 6, 4)` for the first team and `c(2, 3, 5)` for the second team. 

Below is a draft version of the function `get_runner_ranks()`. However, there is an error somewhere. Use any method we discussed in class to identify the error.

```{r, error = T, echo = T, eval = T}
get_runner_ranks <- function(x, y) {
  combined_times <- c(x, y) 
  
  
  sort(combined_times, decreasing = T)
  

  ranks_x <- numeric(length(x))
  ranks_y <- numeric(length(y))
  
  for (i in seq_along(ranks_x)) {
    ranks_x[i] <- match(x[i], combined_times)
  }
  
  for (i in seq_along(ranks_y)) {
    ranks_y[i] <- match(y[i], combined_times)
  }
  

  return(list(x = ranks_x, y = ranks_y))
}
```

5. Below, write a corrected version of `get_runner_ranks()` and compute `get_runner_ranks(c(16.8, 21.2, 19.1), c(17.2, 18.1, 20.0))`.

```{r, error = T, echo = T, eval = T}
get_runner_ranks <- function(team1_times, team2_times) {

  all_times <- c(team1_times, team2_times)
  
  
  all_ranks <- rank(all_times)
  
  team1_ranks <- all_ranks[1:length(team1_times)]
  team2_ranks <- all_ranks[(length(team1_times) + 1):length(all_times)]
  
  result <- list(Team1 = team1_ranks, Team2 = team2_ranks)
  return(result)
}

team1_times <- c(16.8, 21.2, 19.1)
team2_times <- c(17.2, 18.1, 20.0)

ranks <- get_runner_ranks(team1_times, team2_times)
print(ranks$Team1)  
print(ranks$Team2)  

```

6. Explain in your own words what the error was.

The error in the provided function is that the sort function is not saving the sorted vector. 
The sort function returns the sorted vector, but it does not modify the original vector in place. 
To fix the function, you should assign the result of sort to combined_times. 